---
title: "Tarea 2"
Name: "Francisco Vázquez Mendoza"
date: "April"
output: 
  html_document:
    toc: true
    toc_depth: 1
    theme: cerulean
---


<style>
body {
text-align: justify}
</style>

<style type="text/css">
body{/*Normal*/
     font-size: 20px;
}
</style>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr) # A companion package for the book Introduction to Discrete Choice Analysis with `R`
library(mlogit) # Multinomial Logit Models
library(readr) # Read Rectangular Text Data

library(gplots) # Various R Programming Tools for Plotting Data
library(ggmosaic) # Mosaic Plots in the 'ggplot2' Framework
library(treemapify)
library(ggridges)
library(ggalluvial)
library(plotly)
```


Call data set Car from package mlogit. Once you have loaded the data set, do the following:

* How many variables are there in this data set and of which type (i.e., categorical/quantitative)?

* Choose four relevant categorical variables from this data set and describe them using univariate, bivariate, and multivariate techniques. Discuss your results.

* What ideas about individuals’ choices regarding car do you develop from this EDA exercise?


Se carga el conjunto de datos, se muestran las variables y su tipo:

```{r}
data(Car)
glimpse(Car)
```

```{r}
str(Car)
```



En este caso es necesario añadir columnas acerca de la elección, las gráficas se realizarán con la data frame `new_car`:


```{r}
new_car <- Car
choice_char <- gsub("choice","", new_car$choice)
new_car <- new_car |> mutate(choice_char = choice_char)

type <- vector("character", nrow(new_car))
fuel <- vector("character", nrow(new_car))
price <- vector("numeric", nrow(new_car))
range <- vector("numeric", nrow(new_car))
acc <- vector("numeric", nrow(new_car))
speed <- vector("numeric", nrow(new_car))
pollution <- vector("numeric", nrow(new_car))
size <- vector("numeric", nrow(new_car))
space <- vector("numeric", nrow(new_car))
cost <- vector("numeric", nrow(new_car))
station <- vector("numeric", nrow(new_car))
for (i in 1:nrow(new_car)) {
  type[i] <- as.character(new_car[i, paste0("type", new_car$choice_char[i], collapse = "")])
  fuel[i] <- as.character(new_car[i, paste0("fuel", new_car$choice_char[i], collapse = "")])
  price[i] <- new_car[i, paste0("price", new_car$choice_char[i], collapse = "")]
  range[i] <- new_car[i, paste0("range", new_car$choice_char[i], collapse = "")]
  acc[i] <- new_car[i, paste0("acc", new_car$choice_char[i], collapse = "")]
  speed[i] <- new_car[i, paste0("speed", new_car$choice_char[i], collapse = "")]
  pollution[i] <- new_car[i, paste0("pollution", new_car$choice_char[i], collapse = "")]
  size[i] <- new_car[i, paste0("size", new_car$choice_char[i], collapse = "")]
  space[i] <- new_car[i, paste0("space", new_car$choice_char[i], collapse = "")]
  cost[i] <- new_car[i, paste0("cost", new_car$choice_char[i], collapse = "")]
  station[i] <- new_car[i, paste0("station", new_car$choice_char[i], collapse = "")]
}

new_car <- new_car |> mutate(choice_num= factor(choice_char), type=type, fuel=fuel, price=price, range=range,
                             acc=acc, speed=speed, pollution=pollution, size=size, space=space, cost=cost,
                             station=station)

new_car$type <- factor(new_car$type, levels = levels(new_car$type1))
new_car$fuel <- factor(new_car$fuel, levels = levels(new_car$fuel1))
```


```{r}
summary(new_car)
```

**Se eligen 4 variables categóricas relevantes y se describen utilizando técnicas univariantes, bivariantes y multivariantes. Discuta sus resultados.**

*Univariado*
*Choice*

```{r}
new_car %>% group_by(choice_num) %>% summarise(n = n()) %>% ggplot(aes(fill = choice_num, area = n)) + geom_treemap() + labs(title = "Frequency of choice", fill = "Choice")
```
Discusión: Esta distribución muestra que las 3 opciones preferidas por los usuarios son (por orden de frecuencia): la 5, 3 y 1.



```{r}
new_car %>% group_by(type) %>% summarise(n = n()) %>% ggplot(aes(fill = type, area = n)) + geom_treemap() + labs(title = "Frequency of vehicle type", fill = "Type")
```



```{r}
new_car %>% group_by(fuel) %>% summarise(n = n()) %>% ggplot(aes(fill = fuel, area = n)) + geom_treemap() + labs(title = "Frequency of vehicle fuel", fill = "Fuel")
```



```{r}
new_car %>% group_by(cost) %>% summarise(n = n()) %>% ggplot(aes(fill = cost, area = n)) + geom_treemap() + labs(title = "Cost per mile of travel", fill = "Cost")
```

*Técnica bivariante*


```{r}

new_car %>% ggplot(aes(x = type, y = choice_num, fill = choice)) +
  geom_density_ridges(jittered_points = TRUE,
                      bandwidth = 3.5,
                      position = position_points_jitter(width = 0.05, height = 0),
                      point_shape = '|', 
                      point_size = 3,
                      point_alpha = 1,
                      alpha = 0.7) +
  labs(y = "Choice", x = "Type of vehicle", fill = "Choice")

```

Alternativa de gráfico (más adecuado?)

```{r}

new_car %>%
ggplot(aes(x = type,
fill = choice_num)) +
geom_bar(position = "fill") +
labs(y = "Proportion",
x = "Type of vehicle",
fill="Choice")

```

*Técnica bivariante*


```{r}

new_car %>%
ggplot(aes(x = fuel,
fill = choice_num)) +
geom_bar(position = "fill") +
labs(y = "Proportion",
x = "Fuel",
fill="Choice")

```

*Técnica bivariante*
*Choice - Cost*

```{r}

new_car %>%
ggplot(aes(x = cost,
fill = choice_num)) +
geom_bar(position = "fill") +
labs(y = "Proportion",
x = "Cost",
fill="Choice")

```

*Técnica multivariante*


```{r}

# data preparation
new_car_alluvia <- new_car %>%
  select(fuel, type, choice_num) %>%
  group_by(fuel, type, choice_num) %>%
  summarize(frequency = n(),
            .groups = "drop")

```


```{r}

new_car_alluvia %>% 
  ggplot(aes(y = frequency, 
             axis1 = type,
             axis2 = fuel,
             axis4 = choice_num)) +
  geom_alluvium(aes(fill = fuel),
                width = 1/12,
                color = "black") +
  geom_stratum(width = 1/3,
               fill = "black",
               color = "yellow") +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            color = "green",
            size = 3) +
  scale_x_discrete(limits = c("Fuel", "Type",
                              "Choice"),
                   expand = c(.05, .05)) +
  scale_fill_brewer(type = "qual",
                    palette = "Set1") +
  theme_minimal()

```
